Paper template for {ICPhS} 2023 Prague

### 2024-04-25 manuscript/main.tex

@@

>>>>

\subsection{The present study}

\textbf{INTRO}. Accordingly, the research questions and hypotheses are
as follows: 1. What covert attitudes are linked to the perception
Catalan intervocalic /s/ as either {[}s{]} or {[}z{]}? 2. What is the
relationship between individual social factors (mother tongue, province
of origin) and covert attitudes towards Catalan intervocalic /s/?

@@

\section{Results}

The statistical analysis did not reveal a significant effect of phoneme
on the evaluation of the speaker (p \textgreater{} .05) for any of the
11 evaluative statements. However, the model output did show a
significant association between mother tongue and province for some of
the statements. In what follows, we will explore the results for those
statements.

Participant responses to the statements \emph{Parla bé català} `S/he
speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
origin' showed similar results, as illustrated in Figures \ref{fig:pbc}
and \{fig:oc\}. Those who have both Catalan and Spanish as their L1
rated the speaker more positively regardless of phoneme, compared to
those whose L1 is either Catalan or Spanish.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/pbc.png}
\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
\end{center}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/oc.png}
\caption{Ratings of "És d'origen català" as a function of mother tongue}\label{fig:oc}
\end{center}
\end{figure}

Participant responses to the statements \emph{És simpàtic} `S/he is
nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
results, as illustrated in Figures \ref{fig:simpatic} and \{fig:fiar\}.
Those who are from and live in the province of Barcelona give higher
ratings to the speaker, compared to those who are from and live in
Girona.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/simpatic.png}
\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
\end{center}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/fiar.png}
\caption{Ratings of "És de fiar" as a function of mother tongue}\label{fig:fiar}
\end{center}
\end{figure}

\section{Discussion and Conclusions}

The aim of this exploration study was to investigate whether phoneme
production ({[}s{]} vs {[}z{]}), mother tongue (Catalan vs Spanish vs
both), and province of origin (Girona vs Barcelona) influenced the
attitudes of the participants towards a Catalan native speaker. The
results suggest that these predictors have a different effect on the
ratings of the speaker depending on the social characteristic under
evaluation.

Contrary to what we expected, phoneme was not a significant predictor.
One explanation for this divergent pattern could be that either
participants could not distinguish between {[}s{]} and {[}z{]} or, if
they could, this difference in production was not changing their
attitudes towards the speaker. This finding is interesting considering
that, in our stimuli, the production of the voiceless alveolar fricative
in intervocalic position has been considered a result of Catalan in
contact with Spanish \textbf{(Benet et al., 2012; Davidson, 2020;
Recasens Vives, 1996, p.~268)}, and, from a prescriptivist point of
view, it should be avoided \textbf{(Institut d'Estudis Catalans, 2022)}.
Despite the lack of significance of this predictor, it should be
mentioned that the speaker received more positive ratings when the
{[}z{]} was being produced. Although at this point we can only
hypothesize, it could be that the combination of the phoneme /s/ with
the precedent vowel, might give the listener more cues and have a
stronger social meaning. Future studies should address this issue in
order to explore if the combination of vowel + /s/ carries other social
connotations. \textbf{REVISAR}
